\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-10 analysis/paper/paper.Rmd
working on intro text

Mapping forests is crucial to understanding the effects and feedbacks of changes in climate, land use, and disturbance events across diverse landscapes [@saladin2020rapid; @grossiord2020having; @fassnacht2016review]. 
An important aspect of forest mapping efforts is species composition, which describes the amount and distribution of plant species across landscapes. 
Species composition has been shown to be a fundamental metric for evaluating biodiversity [@williams2020remote], identifying habitats [@nagendra2013remote], providing essential inputs to many ecological models [@pecchi2019species], and informing conservation and restoration efforts [@camarretta2020monitoring]. 
Remote sensing technologies enable us to estimate species composition more effectively than using field-based approaches alone. 
We currently have unprecedented access to publicly available hyperspectral airborne imagery at sites around the United States from the National Ecological Observatory Network (NEON) [@keller2008continental]. 
The NEON Airborne Observation Platform (AOP) payload consists of an imaging spectrometer, high resolution digital camera, and light detection and ranging (LiDAR) sensors to capture both spectral and structural measurements of ecosystems [@kampe2010advances].

The National Ecological Observatory Network (NEON) is a valuable source of publicly available ecological data across the United States [@keller2008continental].
We have extracted all NEON AOP hyperspectral data for every mapped stem in NEON plots where field data and airborne remote sensing data were collected during the same year.

We used a machine learning approach to evaluate plant identification potential using the hyperspectral data at multiple taxonomic resolutions: species and genus levels.

Species classification is an active research area [@scholl2020integrating].

>>>>

Mapping forests is crucial to understanding the effects and feedbacks of changes in climate, land use, and disturbance events across diverse landscapes [@saladin2020rapid; @grossiord2020having; @fassnacht2016review]. 
An important aspect of forest mapping efforts is species composition, which describes the amount and distribution of plant species across landscapes. 
Species composition has been shown to be a fundamental metric for evaluating biodiversity [@williams2020remote], identifying habitats [@nagendra2013remote], providing essential inputs to many ecological models [@pecchi2019species], and informing conservation and restoration efforts [@camarretta2020monitoring]. 
Although remote sensing technologies enable us to estimate species composition more effectively than using field-based approaches alone, species classification is still a challenging task.

We currently have unprecedented access to publicly available hyperspectral airborne imagery at sites around the United States from the National Ecological Observatory Network (NEON) [@keller2008continental]. 
NEON-monitored sites are spread across ecoclimatic domains with a diverse array of ecosystems and land cover types including deciduous and evergreen forests, shrub/scrub and woody wetlands, mixed forest and grasslands. 
Field technicians collect field-based vegetation measurements including species observations on a semi-annual basis. 
In addition, NEON’s Airborne Observation Platform (AOP) payload carries a suite of sensors to collect both spectral and structural measurements of ecosystems [@kampe2010advances]. 
The AOP includes an imaging spectrometer which captures hyperspectral reflectance data in over 400 bands across the visible and near infrared wavelengths. 
Hyperspectral imagery is frequently used to classify species with accuracies ranging widely anywhere from 45% to 98% [@fassnacht2016review]. 
Hyperspectral data are capable of capturing reflectance variations and features with hundreds of spectral bands better than multispectral remotely sensed data with a handful to tens of bands [@awad2018forest].
Many factors influence tree species classification results including the number of species being discerned, the homogeneity of species, the density and structural complexity of the forest, and the taxonomic level of classification [@sothe2019tree]. 
Methods are often performed on small scales or in simple ecosystems, and are not successful at scaling to larger areas or different forest types **(cite)**.
Although the demand for species mapping is justified in many cases, there is also potential for both using and finding value in classifying individuals at other taxonomic levels, as explored in a recent bioacoustics study [@cramer2020chirping]. 
In the plant realm, identifying taxons at the genus level had been shown to be easier than at the species level [@tuominen2018assessment].
[@surovy2019acquisition] report that species are more successfully classified into broader categories such as spruce, pine, and broadleaf, compared to identifying individual species in mixed arrangements.

NEON provides a unique opportunity to test, develop, and tune tree classification methods over a continental scale and at multiple taxonomic resolutions. 
We have extracted all NEON AOP hyperspectral data for every mapped stem in NEON plots where field data and airborne remote sensing data were collected during the same year.
We used a machine learning approach to evaluate plant identification potential using the hyperspectral data at multiple taxonomic resolutions: species and genus levels. 
Our study contributes to the active field of tree species classification research using freely available NEON data and open, reproducible code [@marconi2019data; @scholl2020integrating]
